Supplemental Materials for ``Universals of word order reflect optimization of grammars for efficient communication''

### 2019-10-08 writeup/pnas-word-order-si.tex

We extended the parameter space of grammars by continuously interpolating between grammars, making efficiency a \emph{differentiable} function of grammar parameters.
The parameters of such a \key{differentiable word order grammar} are as follows. 
For each dependency label type $\tau$, we have (1) a \key{Direction Parameter} $a_\tau \in [0,1]$, and (2) a \key{Distance Parameter} $b_\tau \in \mathbb{R}$. 
Each dependent is ordered on the left of its head with probability $a_\tau$ and to the right with probability $1-a_\tau$. 
Then for each set of co-dependents $\{s_1, \dots , s_n\}$ placed on one side of a head, their order outward from the head is determined by iteratively sampling from the distribution $\operatorname{softmax}(b_{\tau_1}, \dots, b_{\tau_n})$ (\cite{goodfellow2016deep}, p. 184) without replacement.

>>>>

We extended the parameter space of grammars by continuously interpolating between grammars, making efficiency a \emph{differentiable} function of grammar parameters.
The parameters of such a \key{differentiable word order grammar} are as follows. 
For each dependency label type $\tau$, we have (1) a \key{Direction Parameter} $a_\tau \in [0,1]$, and (2) a \key{Distance Parameter} $b_\tau \in \mathbb{R}$. 
Each dependent is ordered on the left of its head with probability $a_\tau$ and to the right with probability $1-a_\tau$. 
Then for each set of co-dependents $\{s_1, \dots , s_n\}$ placed on one side of a head, their order from left to right is determined by iteratively sampling from the distribution $\operatorname{softmax}(b_{\tau_1}, \dots, b_{\tau_n})$ (for dependents preceding the head) or $\operatorname{softmax}(-b_{\tau_1}, \dots, -b_{\tau_n})$ (for dependents following the head) (for the definition of $\operatorname{Softmax}$, see \cite[p. 184]{goodfellow2016deep}) without replacement.
